Usability of Modern Display Technologies

### 2014-10-13 mental-model-paper/paper2.tex
add more background and remove placeholder images

\section{Background/ Prior Work/ Literary Review}
There exists a fair amount of previous work pertaining to the advantages as well as the limitations of touch sensitive screens. One limitation that is often discussed in the literature is how typing on flat surfaces with no physical keys to guide the fingers requires heightened visual attention to avoid typos. A study by Hussain Tinwala and Scott MacKenzie suggests that since the visual demand on the user is increased, concentration is diverted from the thoughts being expressed. \cite{Tinwala:2010:ETE:1868914.1868972} Be that as it may, there exist several text entry methods that attempt to address this problem. For instance, Tinwala and MacKenzie present a system that includes auditory and tactile feedback to guide eyes-free entry using speech and non-speech sounds, and by vibrations. \cite{Tinwala:2010:ETE:1868914.1868972} Another proposed solution comes from Tactus Technology, a company in Fremont, California. Employees of Tactus Technology are creating a keyboard with shape-shifting keys that pop up from the screen's surface when needed and recede again when no longer necessary.\cite{Tactus}
\begin{figure}[ht]
\centering
\includegraphics[width=2.7in]{tactus-keyboard.jpg}

\caption{A Tactus keyboard}
\label{figure-sample}
\end{figure}

@@

Touch screens are ubiquitous on tablets and smartphones, but their flat glass surfaces can hinder close reading and accurate typing. People flipping through electronic pages often retain less of what they read than on printed ones, studies suggest. Typing on flat surfaces with no physical keys to guide the fingers requires heightened visual attention to avoid typos, draining the concentration from the thoughts being expressed. Companies are trying to address these problems with new tools adapted from the analog world of three-dimensional typewriter keys, tactile paper pages, and pop quizzes on the blackboard.

Tactus Technology of Fremont, California is developing a keyboard with shape-shifting keys that pop up from the screens surface when needed the reced so that the screen is flat and featureless again. Fluid in tiny microchannels raises the keys up and later recedes, making it appear that they keys have melted away. The technology will be offered later this year as an accessory to the iPad Mini, co-founder Craig Ciesla said. Such tactile features can help build muscle memory and improve accuracy -- skills lost in the rush to touch screes, said Scott MacKenzie, a professor of electrical engineering and computer science at York University in Toronto who specializes in human-computer interaction. "As touchscreen phones lack the tactile feel of a physical keyboard, the visual demand on the user is increased." (Eyes-free Text Entry with Error Correction on Touchscreen Mobile Devices) Many people who type on flat glass screen must keep their eyes focused on the surface to hit the correct key, he said. "It's not just that visual attention is needed, but a lot of visual attention"

This means less focus on the act of composition. Keys that rise and fall don't use up valuable screen real estate when no longer needed, are a good solution. Accurate typing isn't the only problem with touch screes. Many studies suggest that people's memory and comprehension are often better when they read long passages on paper than on screen, said Mariette DiChristina, editor in chief of Scientific American. But electronic textbooks are incorporating ways to compensate for this.

@@

\BibTeX\ supports virtually all kinds of references, including books \cite{dui,sgg,iokit,palmos}, parts of books \cite{userModeLinux}, articles \cite{nielsen:dui-review,heer-shneiderman,stackableThreads,xpkernel}, and conference proceedings \cite{ux-3d,iring,contextFileSearch,osHaskell,hibernator}, to name a few.  If not already included in your \LaTeX\ distribution, download and install the \texttt{url} package to support formatting of URLs; you can usually mention these in the \emph{note} or \emph{howpublished} fields of your \BibTeX\ file.

@@

\begin{figure}
\centering
\includegraphics[width=1in]{space.jpg}

\caption{Another sample figure}
\label{figure-sample2}
\end{figure}

>>>>

\section{Background/ Prior Work/ Literary Review}
There exists a fair amount of previous work pertaining to the advantages as well as the limitations of touch sensitive screens. One limitation that is often discussed in the literature is how typing on flat surfaces with no physical keys to guide the fingers requires heightened visual attention. A study by Hussain Tinwala and Scott MacKenzie suggests that since the visual demand on the user is increased, concentration is diverted from the thoughts being expressed. \cite{Tinwala:2010:ETE:1868914.1868972} There exist several input entry methods, however, that attempt to address this problem. For instance, Tinwala and MacKenzie present a system that includes auditory and tactile feedback to guide eyes-free entry using speech and non-speech sounds. \cite{Tinwala:2010:ETE:1868914.1868972} Another proposed solution comes from Tactus Technology, a company in Fremont, California. Employees of Tactus Technology are creating a keyboard with shape-shifting keys that pop up from the screen's surface when needed and recede again when no longer necessary.\cite{Tactus} Craig Ciesla, the co-founder of the company, said that the keyboard will be offered later this year.

\begin{figure}[ht]
\centering
\includegraphics[width=2.7in]{tactus-keyboard.jpg} 
\caption{A Tactus keyboard}
\label{figure-sample}
\end{figure}

Another limitation of touchscreen devices that is frequently examined in the literature is how the human finger as a pointing device has ``very low resolution."\cite{Albinsson} With mobile touch screens in particular, research has illustrated how it is difficult for users to point at targets that are often smaller than the width of their finger. A notable approach to tackling this problem was made by Schneiderman, Sears, and colleagues.\cite{Sears} Their basic technique provides a cursor above the user's finger tip with a fixed offset when touching the screen. The user drags the cursor to a desired target and then lifts the finger to select the targeted object. Although this technique allows users to perform precise movements with generally less errors, the user's efficiency tends to suffer.

Instead of using a bare finger, in some cases the user may use a stylus (pen) to interact with touch screens. A stylus is much ``sharper" pointer than a finger tip, but its resolution may still not be as good as a mouse cursor. A study shows taht pen input is emerging as a promising interaction modality for touch screens.\cite{Bi} The drawbacks to the stylus solution, as discussed in a study by Forlines and Balakrishnan, include the occlusion of other elements on the screen.\cite{Forlines}

Zooming is another method that has been investigated in attempts to increase precision of touchscreen interaction. It is possible to use zooming to enlarge the information space enough so that the user can comfortably point at the target with a bare finger. A common implementation of this zooming technique is called \textit{bounding box zoom}.\cite{Albinsson} With this approach, the user first activates a zooming mode, and then draws a `bounding box' that encloses the area to be zoomed in. When the user releases the box, the system zooms and pans to the selected are and the mode switches back to pointing. One zooming graphical interface that implements the \textit{bounding box zoom} technique is \textit{Pad++}.\cite{Bederson} Though the \textit{Pad++} helps users effectively touch things on the screen, one cost is that ``designers can no longer rely on the users' familiarity with metaphorical reference and this has learnability consequences." \cite{Bederson}

@@

This means less focus on the act of composition. Keys that rise and fall don't use up valuable screen real estate when no longer needed, are a good solution. Accurate typing isn't the only problem with touch screens. Many studies suggest that people's memory and comprehension are often better when they read long passages on paper than on screen, said Mariette DiChristina, editor in chief of Scientific American. But electronic textbooks are incorporating ways to compensate for this.

@@

@@
